---
title: "Differential Privacy for Mobility Data used in Epidemiological Models"
format: revealjs
editor: visual
params:
  common_start: 22
  common_end: 35
  uncommon_start: 6
  uncommon_end: 15
---

## Background

```{r, echo=FALSE, include=FALSE}
library(here)
source(here("code","src.R"))
city_list <- create_cities()
trans_list <- create_all_trans(city_list)
chosen_trans <-   list(
    "Common" = list("start" = params$common_start, "end" = params$common_end),
    "Uncommon" = list("start" = params$uncommon_start, "end" = params$uncommon_end)
  )
user_trans_list <- add_trans(trans_list, chosen_trans)
```

We begin with a simulated set of "cities". In this case there are 36 cities on a 6x6 grid located 100km from their closest neighbor. This is a toy example and real transition data has varying spatial and temporal aggregations.

## Background

```{r}
plot_cities(city_list)
```

## Background

Notice that bigger cities have more transitions between them and some smaller cities don't have any major transitions outside of those to their neighbors.

## Background

```{r}
plot_trans(city_list, trans_list)
```

## Your transitions

How let's check out the two common and uncommon transitions that you've chosen. Take a moment to look at:

1\) How many other individuals move through your set of transitions?

2\) Even though the data are at the "city" level, would your movement be easily identified?

## Your transitions

```{r}
plot_trans(city_list, user_trans_list)
```

## Aggregation

-   What happens when your data are aggregated?

-   Can you still identify your transitions easily?

## Aggregation

```{r}
#aggregate data
x <- aggregate_network(city_list, user_trans_list)

#visualize aggregated data
plot_trans(x$agg_city, x$agg_trans, agg = T)
```

## Not all transitions are equal

-   Depending on the "uncommon" transition you've chosen, it's likely that even with aggregation, the single trip could be easily identified

-   These transitions can identify individuals, but the behavior they represent may be important for epidemiological models, depending on your question of interest

## Identifiable Transitions

1\) Those that move at different times than others (ex. night shift)

2\) Those that move against the standard flow (counter current commutes)

3\) Those in rural areas that may be a part of a small number

## Let's see what happens when we add a tiny bit of noise

-   What do you notice about the transition network before aggregation? How do the transitions compare to the baseline?

## A little bit of noise...

```{r}
#visualize with small amounts of noise
noisy_trans <- add_noise(user_trans_list, noise = 0.1)
a <- plot_trans(city_list, noisy_trans, noise = 0.1)
x <- aggregate_network(city_list, noisy_trans)
b <- plot_trans(x$agg_city, x$agg_trans, agg = T)
ggarrange(a,b, common.legend = T, ncol = 1, legend = "right")
```

## And a little bit more noise...

```{r}
#visualize with small amounts of noise
noisy_trans <- add_noise(user_trans_list, noise = 0.5)
a <- plot_trans(city_list, noisy_trans, noise = 0.5)
x <- aggregate_network(city_list, noisy_trans)
b <- plot_trans(x$agg_city, x$agg_trans, agg = T)
ggarrange(a,b, common.legend = T, ncol = 1, legend = "right")
```

## And a little bit more noise still...

```{r}
#visualize with small amounts of noise
noisy_trans <- add_noise(user_trans_list, noise = 0.8)
a <- plot_trans(city_list, noisy_trans, noise = 0.8)
x <- aggregate_network(city_list, noisy_trans)
b <- plot_trans(x$agg_city, x$agg_trans, agg = T)
ggarrange(a,b, common.legend = T, ncol = 1, legend = "right")
```

## Follow up questions

-   What happens to the transition network as you add more noise?

-   What happens to the aggregate transitions (common and uncommon) as you add noise?

-   What do you think happens to metrics such as disease spread, if these networks are used to simulate transitions between spatial regions?

## Our Analysis

-   We simulated thousands of epidemiological metapopulation models which are often used to evaluate resource allocation and vaccine distribution

-   We evaluated various common epidemiological parameters (outputs) of interest from these models as we increased the noise included in our mobility data

## Our Analysis

-   We found that most metrics are stable through a wide range of noise (much larger than previous studies have allowed).

-   This enables us to provide **significant privacy protection** to individuals who contribute their data to the analysis **without biasing** the results of analysis

-   This relates directly to trust:

    -   Trust in appropriate stewardship of privacy and identifiable data

    -   Trust in the outputs of epidemiological models that are not biased by privacy preserving mechanisms

## Key Questions

-   Who gets to decide what spatial and temporal scales are to be used?

-   Who gets to decide how much privacy preserving noise is added to the data?

-   Who gets to drive the research questions which inevitably will set the balance point between bias and privacy?

## Conclusion

-   Feel free to try it out again with different common and uncommon transitions!

-   For more information please check out our lightning talk at 10:50AM
